\vspace{24pt} Comparison of Different Data Layouts for Columnar In-Memory Databases on the Basis of an Example Application

### 2018-08-31 paper/paper.tex
fixed error in haversine formula
add benchmark: single time frame

\begin{equation} a = sin^2(\frac{\Delta_{lat}}{2}) + cos(lat_1) \cdot sin^2(\frac{\Delta_{lon}}{2}) \end{equation}
\begin{equation} c = 2 \cdot atan2(\sqrt{a}, \sqrt{1 - a}) \end{equation}
\begin{equation} d = R \cdot c \end{equation}

@@

\subsubsection{Get points from timeframe}

>>>>

\begin{equation} a = sin^2(\frac{\Delta_{lat}}{2}) + cos(lat_1) \cdot cos(lat_2) \cdot sin^2(\frac{\Delta_{lon}}{2}) \end{equation}
\begin{equation} c = 2 \cdot atan2(\sqrt{a}, \sqrt{1 - a}) \end{equation}
\begin{equation} d = R \cdot c \end{equation}

@@

\subsubsection{Get points from timeframe}
For our heatmap is necessary to get all taxis within a specific time frame. With every returned GPS point the heatmap also gets a weight which indicates the number of taxis there. It is also possible to change the granularity of those GPS points. We did not implement this task for the Key-Value format, because the overhead of parsing the large string as JSON object before any further operations can be performed is too much.

For the Points format all entries simply are filtered by their timestamp to be within the given period and, depending on the desired granilarity, rounded to a different amount of decimals. Afterwards the results are grouped by latitude and longitude to get the weight. This takes about 19 milliseconds.

As the Frame format is geared to such time frames it should be much faster to get all points within a frame. With 14 milliseconds it is faster than the Points format, but only by a small amount. The reason for that is the delta encoding for all but the first coordinate within a frame. So for each found entry the coordinate needs to be calculated. After this is done, rounding the coordinates according to the granulatrity and grouping to get a weight is the same as for the Points format.
\begin{figure}[ht]
  \centering
  \includegraphics[width=0.48\textwidth]{img/bench_single_frame.png}
  \caption{Time in ms required for retrieving one single time frame}
  \label{fig:bench_whole_trajectory}
\end{figure}
